---
title: "Homework 2"
author: "Hoang Chu"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, echo=FALSE, message=FALSE}
knitr::opts_chunk$set(comment = NA, message=F, warning=F)
library(faraway)
library(magrittr)
library(tidyverse)
```

# 2.
```{r}
data("uswages")
?uswages
summary(uswages)

m <- lm(wage ~ educ + exper, uswages)
summary(m)
```

For every year of education the model estimates a \$51 per week increase in wage. For every year of experience the model esimates only \$10.

```{r log_wage}
update(m, log(wage, 10) ~ .) %>% summary()
```

Log transformation of the response, means we interpret the coefficients as magitude (multiplicative) changes. So for each year of education the model expects a 4% increase in wage. For each year of experience < 1%.

# 3.

```{r hand_calc}
x <- 1:20
y <- x + rnorm(20)

m <- lm(y ~ I(x^2))
summary(m)

direct_calc <- function(x, y, degree = 2) {
  x_mat <- model.matrix(~I(x^degree))
  solve(crossprod(x_mat), crossprod(x_mat, y))
}

direct_calc(x, y)

map(set_names(3:7), possibly(~ direct_calc(x, y, degree = .), "Error"))
```

# 5.

```{r prostate2}
m <- lm(lcavol ~ lpsa, prostate)
m2 <- lm(lpsa ~ lcavol, prostate)

ggplot(prostate, aes(lcavol, lpsa)) +
  geom_point(alpha = .5) +
  geom_line(aes(x = predict(m), color = "lcavol ~ lpsa")) +
  geom_line(aes(y = predict(m2), color = "lpsa ~ lcavol")) +
  geom_point(aes(y = 2.48, x = 1.36), shape = 1, size = 10)
```

Algebra not showm but you can calculate the intesection if you solve for either `lcavol` or `lpsa` in the system of equations represented by the model coefficients.

# 6.

```{r cheddar}
data("cheddar")
summary(cheddar)
```

## a.

```{r cheddar_a}
m <- lm(taste ~ ., cheddar)
summary(m)
```

## b.

```{r cheddar_b}
cor(m$fitted.values, cheddar$taste)^2
```

R-Squared, the coeffecient of determination or percentage of variance explained

## c. 

```{r cheddar_c}
m2 <- update(m, . ~ -1 + .)
cor(m2$fitted.values, cheddar$taste)^2
```

## d.

```{r cheddar_d}
m_mat <- model.matrix(m)

qr_decomp <- qr(m_mat)

backsolve(
  qr.R(qr_decomp), # upper-right
  t(qr.Q(qr_decomp)) %*% cheddar$taste
)
```